---
title: "Raining In Seattle"
author: "Faris Nour"
date: '2018-01-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem

You're about to get on a plane to Seattle. You want to know if you should bring an umbrella. You call 3 random friends of yours who live there and ask each independently if it's raining. Each of your friends has a 2/3 chance of telling you the truth and a 1/3 chance of messing with you by lying. All 3 friends tell you that "Yes" it is raining. What is the probability that it's actually raining in Seattle? [Link](https://goo.gl/ndJ5oJ)

This document will run through a simulation of this problem and give us an answer based on the average probability. Modifying `S` will run the experiment more times

```{r cars}
S <- 2000
n <- 500
count <- rep(0, S)
positive <- rep(0, S)
for (i in 1:S) {
  for (j in 1:n) {
    raining <- sample(x = c(TRUE, FALSE), prob = c(1/4, 3/4), size = 1)
    friend1 <- sample(x = c("yes", "no"), size = 1, prob = c(2/3, 1/3), replace = TRUE)
    friend2 <- sample(x = c("yes", "no"), size = 1, prob = c(2/3, 1/3), replace = TRUE)
    friend3 <- sample(x = c("yes", "no"), size = 1, prob = c(2/3, 1/3), replace = TRUE)
    if (raining == FALSE) friend1 <- if (friend1 == "yes") "no" else "yes"
    if (raining == FALSE) friend2 <- if (friend2 == "yes") "no" else "yes"
    if (raining == FALSE) friend3 <- if (friend3 == "yes") "no" else "yes"
    if (friend1 == "yes" && friend2 == "yes" && friend3 == "yes") {
      count[i] <- count[i] + 1
      if (raining) positive[i] <- positive[i] + 1
    }
  }
}

sprintf("the average among all experiments is %.5f", mean(positive/count))

```
Compare this with the actual answer which is 8/11 = 0.72727272


## Additional Plots

For each of the S = 2000 experiments, we can generate a histogram 

```{r probability, echo=FALSE}
hist(positive/count, main = "Histogram of estimated probability", xlab = "Estimated Probability with n=500")
abline(v = mean(positive/count), col = "red", lwd = 2)
legend(x = "topright",
  c("Mean"),
  col = c("red"),
  lwd = c(2))
```

We can see that the empirical simulation supports the analytical answer.
